---
title: "eBird Abundance Data"
author: "Peter Hoff"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

### Summary

The file `eBirdAPI.r` provides a few R functions to download and wrangle data from the [eBird project](https://ebird.org/science/download-ebird-data-products), 
using their API. The functions include the following:

* `eBirdGetURL`: Download generic data and leave it in json/list format
* `eBirdGetData` : Download *complete* checklists from a given region on a given date. Returns a list that includes 
  * `obsData`: numeric matrix of bird counts by checklist and species;
  * `metaData`: dataframe of checklist metadata. 
* `eBirdMergeData`: Merge two lists that have the format returned by `eBirdGetData`;
* `eBirdNames` : Not a function, but a table that matches the bird codes used by eBird to common names. 

### Comments:

1) To download data from eBird you need to have an eBird API key, 
and set it as an 
environment variable with `Sys.setenv(EBIRD_KEY = "xxxxxxxx")`.   
2) There are some duplicates (friends who collect and report the same data).
This should be fixed in the future. 
3) Some birders just report presence without a numeric count. For these cases, numeric counts are imputed, but all imputed values end in `.5` so that they can be identified. 
4) The eBird API only allows 200 checklists to be downloaded per request. 
 If this limit is reached then only the first 200 lists are included in the 
 dataset, and a warning message is printed.    
5) There exist a couple of R packages for accessing eBird data: [rebird](https://github.com/ropensci/rebird) and [auk](https://github.com/CornellLabofOrnithology/auk/). The latter assumes you have already downloaded a super-large eBird database. The former uses the API (like the code in this repo), but 
doesn't seem to provide access to information on 
checklist completeness (which is important to know for statistical analysis). 


### How to use 

Load in the functions and see what they are:
```{r}
#source("https://raw.githubusercontent.com/pdhoff/eBirdGetData/master/eBirdAPI.r")

source("eBirdAPI.r")

objects()
```

Download and merge one week of data from two NC counties (Durham and Beaufort):

```{r}
counties<-c("US-NC-063","US-NC-013")
dates<-as.Date(0:6,origin = "2020-05-01")

eBdat<-NULL
for(i in seq_along(counties)){
  for(j in seq_along(dates)){
    eBdat<-eBirdMergeData(eBdat,eBirdGetData(region=counties[i],date=dates[j]))
    cat(counties[i],dates[j],"\n")
}}
```

Examine metaData a bit:
```{r}
eBdat$metaData[1:3,]

plot(eBdat$metaData$long,eBdat$metaData$lat,xlab="longitude",ylab="longitude") 
```

Examine obsData a bit:
```{r}
eBdat$obsData[1:3,1:6] 

## Obtain counts per hour of effort 
obsPH<-sweep(eBdat$obsData,1,eBdat$metaData$duration,"/") 

## Top 20 birds in Durham
sort(apply(obsPH[eBdat$metaData$loc3==counties[1],],2,mean),decreasing=TRUE)[1:20]  

## Top 20 birds in Beaufort 
sort(apply(obsPH[eBdat$metaData$loc3==counties[2],],2,mean),decreasing=TRUE)[1:20]  
```





